Constraining the Interiors of Asteroids Through Close Encounters

### 2022-04-15 paper/appendices.tex
Updated paper

>>>>

\section{Detailed analysis of the dependence of PPD width on encounter parameters}

\subsection{Orbital elements}
\label{sec:scan-orbit}
A Keplerian orbit is completely described by five parameters, but three describe the orbit's orientation with respect to the central body. The orbit can be rotated to the $\unit X \unit Y$-plane, as mandated by our coordinate definitions, by changing the density moments of the central body. Since $J_{00}$ is unchanged by this rotation and $J_{1m}=0$, the orbit orientation is irrelevant until the $J_{2m}$ terms of equation \ref{eqn:tidal-torque}, and we do not investigate them here.

We parametrize the shape of the orbit by the perigee distance $r_p$ and excess velocity $v_\infty$. Fits of the type described in section \ref{sec:fit} were run for many values of $r_p$ and $v_\infty$ and the 1 and 2$\sigma$ confidence intervals are displayed in figures \ref{fig:scan-vex} for $v_\infty$ and \ref{fig:scan-perigee} for $r_p$.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-vex.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of orbital excess velocity $v_\infty$. The vertical dashed line indicates the reference asteroid value of 6 km s$^{-1}$. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-vex}
\end{figure}

Figure \ref{fig:scan-vex} demonstrates that parameter precision does not depend strongly on excess velocity, aside from a slight trend especially in the higher order parameters for uncertainty to increase with $v_\infty$. This is likely due to the fact that larger $v_\infty$ leads to a faster and flatter orbit with less time spent close to the planet, where tidal torque is strongest. There are also smaller-scale oscillations in the uncertainty, due to the orientation of the asteroid at perigee varying. The asteroid is always simulated to start at the same orientation, but increasing $v_\infty$ decreases the time to perigee, so that the asteroid enters this region of high torque at different orientations depending on $v_\infty$. This effect explains why these small-scale oscillations have the same period for all parameters. Note that these oscillations are sometimes large enough to raise $\sigma > 0.01$, shown by the red vertical line.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-perigee.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of perigee distance $r_p$. The vertical dashed line indicates the reference asteroid value of 5 Earth radii. The dotted curve indicates a power-law fit (see text). The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-perigee}
\end{figure}

Figure \ref{fig:scan-perigee} shows much stronger dependence of parameter uncertainty on perigee distance, as expected by the factor of $(a_\mathcal{A}/D)^{\ell'}$ present in equation \ref{eqn:tidal-torque} and mentioned in section \ref{sec:tidal-torque}. For $r_p \approx 6$ or 7 Earth radii, $\sigma(K_{30})$ reaches the reference 0.01 limit. At slightly higher $r_p$, $K_{31}$ and $K_{32}$ become similarly unresolved, and at $r_p \approx 20$ Earth radii, $K_{33}$, the last second-order component, becomes unresolved. Even before this, at $r_p \approx 10$ Earth radii, the most uncertain parameter $K_{30}$ fills the prior distribution with uncertainty ranging from -1 to 1, visible by the sudden cut-off in uncertainty increase and the discontinuity of the $\sigma$ curve there. The location of these limits will change if different encounter properties are used, but this is still an illustrative example.

\begin{table}
  \centering
  \begin{tabular}{l|c}
    \hline
    Parameter & $\alpha$ \\ \hline
    $\gamma_0$ & $2.05$ \\
    $K_{22}$ & $5.47$ \\
    $K_{20}$ & $5.47$ \\ \hline
    $\Re K_{33}$ & $3.35$ \\
    $\Im K_{33}$ & $3.37$ \\
    $\Re K_{32}$ & $3.05$ \\
    $\Im K_{32}$ & $3.27$ \\
    $\Re K_{31}$ & $3.53$ \\
    $\Im K_{31}$ & $4.02$ \\
    $K_{30}$ & $5.75$ \\ \hline
  \end{tabular}
  \caption{Power law slope values for the dependence of parameter uncertainty on perigee distance $r_p$. Slope is defined by $\sigma \propto r_p^\alpha$.}
  \label{tab:scan-perigee-alpha}
\end{table}

Fitted to each of the curves in figure \ref{fig:scan-perigee} are power law uncertainties $\sigma \propto r_p^\alpha$. These fits were performed via the method of least squares, and all data with $\sigma > 0.7$ was removed due to its sensitivity to the arbitrarily-chosen boundary of the prior. The values of $\alpha$ are shown in table \ref{tab:scan-perigee-alpha}. These slope values express how much each parameter is dependent on $r_p$. It is observed that $\gamma_0$ is least dependent on $r_p$, with $\sigma(\gamma_0) \sim r_p^2$. The other two first-order parameters are much more strongly dependent, with $\sigma \sim r_p^{5.5}$. The second-order parameters have milder slopes between 3 and 4 (except for $K_{30}$) which is fortunate from the perspective of making precise observations, since lower $\alpha$ renders larger values of $r_p$ accessible to measuring $K_{3m}$.

The axes of figure \ref{fig:scan-perigee} show that parameters with large $m$ are more precisely determined than parameters with small $m$, as can be seen by comparing $K_{22}$ to $K_{20}$ and comparing $K_{33}$ to other $K_{3m}$ values. Large $m$ moments correspond to moments that control higher frequency fluctuations in density at the asteroid equator. This pattern of large $m$ corresponding to low uncertainty is a general trend and will be seen in the following sections as well.

The very strong dependence of $\sigma$ on $r_p$ makes this analysis only to extract second-order moments on close encounters. Fortunately, in the case of Earth, these encounters are also likely to have the best associated observational uncertainty when above the horizon due to their proximity. The first-order moments can still be extracted at much larger perigee distances in our model.

\subsection{Observational Uncertainty}
\label{sec:scan-uncertainty}
There are two parameters, $\sigma_\theta$ and $\sigma_\rho$, which govern the observational uncertainty of the data set (defined in section \ref{sec:uncertainty}). Rather than explore the full space spanned by these two values, we measure how parameter uncertainties depend on the product of uncertainties $\sigma_\theta\sigma_\rho$ (with the radio fixed), and the ratio $\sigma_\rho / \sigma_\theta$ (with the product fixed).

We choose these metrics because we generally expect that the parameter uncertainty $\sigma$ be proportional to the observational uncertainty, but whether the dependence is stronger on $\sigma_\theta$ or $\sigma_\rho$ is not immediately clear. We get around this problem by varying $\sigma_\theta$ and $\sigma_\rho$ together and fixing their ratio, and measuring the posterior uncertainty $\sigma$, with $\sigma/\sigma_\theta$ shown in figure \ref{fig:scan-product}. Indeed, we find that $\sigma \propto \sigma_\theta$ almost exactly, and since $\sigma_\rho / \sigma_\theta$ is fixed, we also have $\sigma \propto \sigma_\rho$. For large $\sigma_\theta \sigma_\rho$, the proportionality fails, but this is because $\sigma(K_{30}) \approx 1$ which fills the prior. Uncertainty cannot increase beyond this value.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-product.pdf}
  \caption{1 and 2$\sigma$ confidence intervals divided by $\sigma_\theta$ for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty product $\sigma_\theta \sigma_\rho$. The vertical dashed line indicates the reference asteroid value of $10^{-9}$. The red vertical lines indicate when $\sigma =0.01$.}
  \label{fig:scan-product}
\end{figure}

We also investigate the dependence of posterior uncertainty on $\sigma_\rho / \sigma_\theta$ with $\sigma_\theta \sigma_\rho$ fixed in figure \ref{fig:scan-ratio}. If we simply had $\sigma \propto \sigma_\theta \sigma_\rho$, then $\sigma$ would have no dependence on $\sigma_\rho / \sigma_\theta$. Any deviation from constant $\sigma$ shown in the figure therefore reveals some additional dependence in the model on one of the observational uncertainties.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-ratio.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational uncertainty ratios $\sigma_\rho / \sigma_\theta$. The reference asteroid value is $\sigma_\rho/ \sigma_\theta =10^{-5}$. The red vertical lines indicate when $\sigma = 0.01$. (Lines are not shown for $K_{3m}$, $m < 3$ because they coincide with the vertical axis.)}
  \label{fig:scan-ratio}
\end{figure}

Indeed, figure  \ref{fig:scan-ratio} shows increased uncertainty when $\sigma_\rho/\sigma_\theta$ is large, so that $\sigma$ depends more on $\sigma_\rho$ than on $\sigma_\theta$. We summarize this pattern by stating that, if the observer had to choose between better precision on the period or on the spin pole of the data, they should choose period. This is fortunate for observers since, \jtd{I want to say that precision on the period is better constrained by light curve analysis, but I do not know enough about light curve analysis to be sure.}

Another conclusion that can be drawn from figure \ref{fig:scan-ratio} is that $\sigma(\sigma_\rho/\sigma_\theta)$ is proportional for all $K_{\ell m}$. The proportionality constant shows the same dependence on $\ell$ and $m$ mentioned previously. The proportionality is broken near $\sigma \approx \pm 1$, where $\sigma$ fills the prior.

The location of the $\sigma > 0.01$ limit demonstrates the great influence of observational precision on fit uncertainty; increasing $\sigma_\theta \sigma_\rho$ or $\sigma_\rho / \sigma_\theta$ by even a small amount raises $\sigma(K_{3m}) > 0.01$ in most cases.

\subsection{Asteroid shape}
\label{sec:scan-shape}

The true values of $K_{\ell m}$, $\gamma_0$, and $a_\mathcal{A}$ affect the uncertainties in extracted density moments $\sigma$. Here, we only investigate the sensitivity of $\sigma$ to the first-order parameters and $a_\mathcal{A}$. The $K_{2m}$ moments can therefore also be viewed as the axes of a uniform density triaxial ellipsoid (equation \ref{eqn:ellipsoid-axes}).

In figure \ref{fig:scan-space-sigma}, we show the 1$\sigma$ confidence intervals as a function of $K_{20}$ and $K_{22}$, or alternatively $a/c$ and $b/c$. We use axis ratios rather than the values of $a$, $b$, and $c$ to remove the $a_\mathcal{A}$ dependence of equation \ref{eqn:ellipsoid-axes}. The figure shows large uncertainty in $\gamma_0$ for $K_{22}=0$, or $a/c=b/c$, because $K_{20}$ is rotationally symmetric around $\unit z$, and $\gamma_0$ is the initial orientation with respect to the $\unit z$ axis. The data then have no physical dependence on $\gamma_0$ when $K_{22}=0$. This induces degeneracy in the model which inflates uncertainties, not only in $\gamma_0$ but also the other components.

\begin{figure*}
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-1-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-2-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-theta-3-sigma.pdf}

  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-1-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-a-sigma.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-b-sigma.pdf}

  \caption{1$\sigma$ confidence interval for fit parameters $\gamma_0$, $K_{22}$, and $K_{20}$ (\textit{top row}) and $\gamma_0$, $a/c$, and $b/c$ (\textit{bottom row}). Also shown as black points are the reference asteroid shapes; the symmetric case is marked with a red circle and the asymmetric with a black triangle.}
  \label{fig:scan-space-sigma}
\end{figure*}

To remove the inflated uncertainty, one could assume a rotationally symmetric asteroid, remove $\gamma_0$ as a parameter, and run a fit. For a nearly rotationally symmetric asteroid however, a new parametrization is necessary which does not contain the ill-constrained $\gamma_0$ parameter. This task is beyond the scope of this paper, so we mostly consider asymmetric asteroids throughout.

Figure \ref{fig:scan-space-sigma} also shows low uncertainty for highly asymmetric asteroids, where $b/c$ and $a/c$ are very different (i.e., when $|K_{22}|$ is large). Additionally, $\sigma(K_{20})$ and $\sigma(K_{22})$ decrease for large $|K_{20}|$, which corresponds to large axis ratios in the ellipsoid case.

Figure \ref{fig:scan-space-corr} displays the correlation between the first-order parameters for reference. They show that $\gamma_0$ and $K_{22}$ are often correlated for asymmetric asteroids, while $\gamma_0$ and $K_{20}$ are usually not. This is expected as $K_{22}$ is dependent on the orientation of the asteroid and $K_{20}$ is not. They also show that $K_{22}$ and $K_{20}$ are usually correlated, and that $a/c$ and $b/c$ are highly correlated. The latter is expected due to the $1/c$ dependence. As for the former, this correlation could likely be removed by an alternate parametrization, reducing uncertainties in the shape parameters.

\begin{figure*}
  \includegraphics[width=0.33\textwidth]{figs/probe-space-corr12.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-corr13.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-corr23.pdf}

  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-1b.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-1a.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/probe-space-ab-ab.pdf}

  \caption{Correlations between PPDs for fit parameters $\gamma_0$, $K_{22}$, and $K_{20}$ (\textit{top row}) and $\gamma_0$, $a/c$, and $b/c$ (\textit{bottom row}).  Also shown as black points are the reference asteroid shapes; the symmetric case is marked with a red circle and the asymmetric with a black triangle.}
  \label{fig:scan-space-corr}
\end{figure*}

Overall, the variation in the uncertainties on $K_{20}$ and $K_{22}$ (the first-order density moments) is present but largely smooth across their allowed parameter space, as is their correlation (except the large $K_{22}$ corner). It therefore seems reasonable to use the asymmetric asteroid shape as a stand-in for an unknown's asteroid shape when simulating an encounter, as we do in this paper. The uncertainty then can be expected to differ across other shapes by a factor of about two or less, as long as the degenerate, symmetric asteroid regime is avoided.

On the other hand, the posterior uncertainty of $K_{3m}$ is much more strongly dependent on asteroid length $a_\mathcal{A}$. Figure \ref{fig:scan-am} displays posterior uncertainty $\sigma$ as a function of $a_\mathcal{A}$, defined in equation \ref{eqn:am}. The vertical axis is shown in log space, meaning that the average of the upper and lower error bars of $\sigma$ is shown instead of showing both as is done in other figures.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-am.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of asteroid length $a_\mathcal{A}$. The reference cadence is 2 minutes. The red vertical lines indicate when $\sigma = 0.01$. The black dashed line indicates the reference value of $a_\mathcal{A}=1000$ m.}
  \label{fig:scan-am}
\end{figure}

As was mentioned in section \ref{sec:tidal-torque}, the $K_{2m}$ parameters are insensitive to $a_\mathcal{A}$ since the $a_\mathcal{A}^2$ term in $\bm \tau$ (equation \ref{eqn:tidal-torque}) is canceled by the $a_\mathcal{A}^2$ in the moment of inertia (equation \ref{eqn:moi}). The $K_{3m}$ uncertainty is strongly dependent on $a_\mathcal{A}$ for the same reason that uncertainty is strongly dependent on $r_p$: the $(a_\mathcal{A}/D)^{\ell'}$ dependence of equation \ref{eqn:tidal-torque}. At $a_\mathcal{A} \lesssim 700$ m, $K_{30}$ becomes unresolved. The other parameters become unresolved for $a_\mathcal{A} \lesssim 300$ m, with the cut-off for $K_{33}$ being less than 100 meters. This $a_\mathcal{A}$ cut-off is quite sharp; figure \ref{fig:scan-am} shows steep decrease in $\sigma(a_\mathcal{A})$ for low $a_\mathcal{A}$, with an approximate functional form of $\sigma(a_\mathcal{A}) \sim e^{(a_\mathcal{A}/C)^p}$ for a scaling constant $C$ and some power $p < 1$. (In figure \ref{fig:scan-am}, the best fitting values are $p < 0.1$.) This sharpness indicates that the cut-off value of $a_\mathcal{A}$ is unlikely to change significantly if other parameters of the flyby are altered in a way that slightly increases or decreases $\sigma$. It therefore appears that extracting precise $K_{30}$ from asteroids with $a_\mathcal{A} \lesssim 700$ m is very unlikely with this analysis, and for $a_\mathcal{A} \lesssim 50$ m, all $K_{3m}$ will likely be unresolved.

For uniform density asteroids, large $a_\mathcal{A}$ is equivalent to large asteroid radius, which was mentioned in section \ref{sec:moments} and explicitly shown in equation \ref{eqn:ellipsoid-axes}. In non-uniform density asteroids, large $a_\mathcal{A}$ can also be achieved by distributing the mass of the asteroid near the surface, because the $r^2$ term in the integrand of the definition of $a_\mathcal{A}$ causes the density of regions distant from the asteroid centre of mass to dominate $a_\mathcal{A}$.

\subsection{Cadence}
\label{sec:scan-cadence}

The time between observations of asteroid angular velocity, or cadence, may vary depending on the observational schedule of the observing telescopes and the path of the asteroid through the sky.  We measure how the posterior uncertainty $\sigma$ varies with cadence ranging from two minutes to one hour in figure \ref{fig:scan-cadence}.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-cadence.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of observational cadence $\Delta t$ The reference cadence is 2 minutes. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-cadence}
\end{figure}

Figure \ref{fig:scan-cadence} displays little dependence of uncertainty on cadence $\Delta t$ for $\Delta t \lesssim 40$ min. We also see flaring of uncertainty for very large cadence, largely driven by the paucity of data points. However, uncertainty dramatically increases for many parameters at about $\Delta t = 30-40$ min, a time scale which is likely characteristic of the asteroid system. We name this rough cadence limit $T_\text{cad}$.

We expect $T_\text{cad}$ to be a function of two dynamical time scales of the system: the rotational period of the asteroid $P_\omega$ and the time spent near perigee $T_p$. The latter can be estimated by
\begin{equation}
  T_p \sim \frac{r_p}{v_\infty}\brackets{2\frac{\mu_\mathcal{B}}{r_pv_\infty^2}+1}^{-\frac{1}{2}}
  \label{eqn:tp}
\end{equation}
which is the ratio of the perigee radius to velocity at perigee. The exact choice of the formula of $T_p$ not obvious, and alternatives to equation \ref{eqn:tp} are possible. For the simulated asteroid, $P_\omega = 9$ hr and $T_p = 42$ min by this definition.

We further study the effect of $P_\omega$ and $T_p$ on $T_\text{cad}$ in appendix \ref{app:cadence-tests}, and find that both affect $T_\text{cad}$ by roughly the same amount.

Figure \ref{fig:scan-cadence} shows that as long as $\Delta t < T_\text{cad}$ is achieved, the influence of cadence on $\sigma$ is minimal (except for $K_{30}$ where the $\sigma<0.01$ limit is slightly exceeded). However, shorter cadence almost always leads to lower uncertainties.

\subsection{Perigee gap}
\label{sec:scan-gap}
In certain circumstances, spin data might not be able to be captured for a close encounter at perigee. The asteroid might dip below the horizon, or it might pass too close to the sun to be observed. Generally, angular velocity data can be collected when the asteroid is distant from the central body, where torque is low. There, the angular velocity evolution is dominated by torque-free precession dictated by the moment of inertia components. That zero-torque data can still be used to fix $K_{20}$ and $K_{22}$ as in \cite{MOSKOVITZ2020113519}. However, $K_{3m}$ are not extractable from precession data alone. We are therefore curious as to how our posterior uncertainties change due to lack of data during the encounter perigee.

To test this, we mask the perigee of the counter by removing a duration $T$ of data centred on the perigee, where $T$ ranges from 0 to 3 hours. To prevent lack of precision on $K_{\ell m}$ induced by lower amounts of data for high $T$, we always cut 3 hr$-T$ from the data set, half from the beginning and half from the end, so that each data set produced for all $T$ has the same length of data before and after the perigee. We then fit the same asteroid model to the cut data for all $T$ and plot posterior uncertainties $\sigma$ in figure \ref{fig:observation-gap}.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/observation-gap.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of a data gap of width $T$ at perigee. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:observation-gap}
\end{figure}

Since torque is greatest at perigee, we expect that region of the data to contain the most information about $K_{\ell m}$, and therefore uncertainty should increase monotonically with $T$, which is seen in figure \ref{fig:observation-gap}. We also see that the first-order parameters are not as sensitive to $T$ as the second-order parameters, because $K_{2m}$ are additionally constrained by torque-free precession after perigee.

Most parameters show dramatically increased uncertainty in the $T \sim 1-2$ hr range. On the other hand, none of the uncertainties increase noticeably for $T < 30$ min. Thirty minutes of dropped data is equivalent to fifteen dropped points for the simulated cadence of $\Delta t = 2$ minutes, showing that many data points can dropped from the data set at perigee before the uncertainty starts to increase.

Qualitatively, \ref{fig:observation-gap} shows similar dependence of $\sigma$ on $T$ as \ref{fig:scan-cadence} showed for $\sigma$ on cadence $\Delta t$. They also both have cut-offs where uncertainty markedly increases, and both the $T$ and $\Delta t$ cut-offs have qualitatively similar shapes although they occur at different values of $\Delta t$ and $T$. This suggests that the factors that govern uncertainty due to cadence (appendix \ref{app:cadence-tests}) also may govern sensitivity to lack of data at perigee in a similar way.

\subsection{Initial spin pole}
\label{sec:scan-spin}

The tidal torque experienced by the asteroid is affected by the initial direction of asteroid spin $\bm \Omega_0$ both because spin sets the initial asteroid orientation up to $\gamma_0$ and because of the spin-dependence of the rotational equations of motion (equation \ref{eqn:omega-eom}).

In figure \ref{fig:scan-spin}, we display 1$\sigma$ uncertainties for all parameters as a function of the direction of $\bm \Omega_0$, mapped onto the unit sphere in the inertial frame. Our samples for $\bm \Omega_0$ were laid out on a Fibonacci sphere to ensure they were roughly evenly spaced (marked in figure \ref{fig:scan-spin-avg}). To highlight common features across the parameters, we also display the average 1$\sigma$ sensitivity in figure \ref{fig:scan-spin-avg}. The average is weighted such that the uncertainty map for each parameter contributes an equal amount
(the weight of each map is set to one-tenth of the map's mean). This average map is presented in two different projections to allow data at $\unit Z$ to be read.

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{figs/spin-pole.pdf}
  \caption{$1\sigma$ uncertainties for the first-order parameters (\textit{top}) and second-order (\textit{bottom}) as a function of the initial direction of spin in the inertial frame. All maps are made in the Mollweide projection. The orange star indicates the reference spin pole. The red contours enclose regions where $\sigma \geq 0.01$.}
  \label{fig:scan-spin}
\end{figure*}

\begin{figure*}
  \centering
  \includegraphics[width=0.45\textwidth]{figs/spin-pole-avg-mollweide.pdf}
  \includegraphics[width=0.45\textwidth]{figs/spin-pole-avg-polar.pdf}
  \caption{The weighted average of the uncertainties shown in figure \ref{fig:scan-spin}, in Mollweide (\textit{left}) and polar (\textit{right}) projections in the inertial frame. See text for a description of how the average was computed. Black dots indicate the Fibonacci-sphere-distributed locations of sample spin poles, and the orange star indicates the reference spin pole. The polar projection is centred at the south pole, or $-\unit Z$.}
  \label{fig:scan-spin-avg}
\end{figure*}

Certain alignments of the body-fixed frame to the inertial frame lead to special conditions on torque, as discussed in section \ref{sec:tidal-torque}. For example, $\bm z \parallel \unit Z$ and $\bm z \parallel \unit Y$ at perigee lead to $\bm \tau \parallel \unit z$ to first-order, and $\bm \tau \parallel \unit X$ at perigee leads to $\bm \tau = 0$ to first-order. We relate this to the initial direction of $\bm \Omega_0$, via the approximation that $\bm \tau$ is small until perigee. Then, since $\bm \omega_0 \parallel \unit z$ in the body-fixed frame as an initial condition (section \ref{sec:sim}), we have that $\bm \Omega_0 \parallel \unit Y$ and $\bm \Omega_0 \parallel \unit Z$ both lead to $\bm \tau \parallel \unit z$, and $\bm \Omega_0 \parallel \unit X$ leads to $\bm \tau = 0$.

Figure \ref{fig:scan-spin-avg} shows area of increased uncertainty for $\bm \Omega_0 \parallel \unit Z$ and $\bm \Omega_0 \parallel \unit Y$, but not the $\unit X$ case. This indicates that $\bm \tau \parallel \unit z$ causes increased uncertainty. Physically, $\bm \tau \parallel \unit z$ only changes an asteroid's rotational period and does not cause it to tumble, eliminating the ability to discern moment of inertia ratios from zero-torque precession after the encounter. Also, $\tau_z$ affected by fewer parameters than $\tau_x$ or $\tau_y$.  Both of these are reasons why $\bm \tau \parallel \unit z$ might inhibit precise fits to spin data. If $\bm \tau = 0$ to first-order, then second-order $\bm \tau$ and non-perigee $\bm \tau$ will dominate, which may increase precision to these usually non-dominant parameters and therefore not have the same increasing effect on $\sigma$.

It is important to note, however, that uncertainty does not vary by much more than a factor of two outside the imprecise regions of $\bm \Omega_0 \parallel \unit Z$ and $\bm \Omega_0 \parallel \unit Y$, though these regions are wide for some parameters. Within the imprecise regions, uncertainty can grow up to four times or more the uncertainty at other $\bm \Omega_0$ values, and can exceed the $\sigma ~0.01$ benchmark. The trends for are roughly consistent across parameters (figure \ref{fig:scan-spin}), leading to clearly visible imprecise regions in the average $\sigma$ (figure \ref{fig:scan-spin-avg}).

\subsection{Rotational period}
\label{sec:scan-period}

We also study the effect of the initial rotational period of the asteroid $P_\omega$ on posterior uncertainty $\sigma$. The dynamical time scales $r_p/v_\infty$ and $\mu_\mathcal{B} / (r_p v_\infty)$ have already been mentioned in the context of the cadence cut-off (section \ref{sec:scan-cadence}), and the ratio between them and $P_\omega$ in principal may affect $\sigma$. In figure \ref{fig:scan-period}, we show $\sigma$ as a function of $P_\omega$ for a range of periods typical of NEOs.

\begin{figure}
  \centering
  \includegraphics[height=0.89\textheight]{figs/scan-period.pdf}
  \caption{1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs (\textit{top}) and second-order parameters (\textit{bottom}) as a function of initial rotational period $P_\omega$. The reference value of 9 hr is shown as a vertical dotted line. The red vertical lines indicate when $\sigma = 0.01$.}
  \label{fig:scan-period}
\end{figure}

Like figure \ref{fig:scan-vex}, depicting the dependence of $\sigma$ on $v_\infty$, figure \ref{fig:scan-period} shows small-scale variation in uncertainty due to the fact that varying the initial period changes the value of $\gamma$ at perigee, which affects uncertainty to a factor of about two. But a large-scale trend is also visible in many parameters. $K_{20}$ and $K_{22}$ show very large uncertainty for $P_\omega \lesssim 4$ hr because these fast-rotators tumble very little after perigee. This increases uncertainty on the $K_{2m}$ parameters, which are constrained by tumbling.

We expect that quickly rotating asteroids would not tumble because, for small $P_\omega$, the dynamical variables $\bm D$, $\bm \omega$, $\alpha$, and $\beta$ vary much smaller than $\gamma$. Approximating each variable as constant over one full rotation of $\gamma$, we can integrate the first-order contribution of $\bm \tau$ over $\gamma \in (0, 2\pi)$ and reveal that each rotation has zero average first-order torque. Since the first-order torque over each rotational period cancels out, there is no secular torque to force the asteroid to tumble. However, this effect does not apply to the second-order parameters, since the integral over the second-order term of $\bm \tau$ does not vanish, as seen in the figure.

Another feature of figure \ref{fig:scan-period} is that $K_{\ell 0}$ is more uncertain at low $P_\omega$ than the other parameters. This is most visible in the figure for $K_{30}$. The cause is likely that $K_{\ell 0}$ cannot contribute to $\tau_z$ as shown in equation \ref{eqn:tidal-torque}. We already discussed that asteroids with small $P_\omega$ do not tumble, and since $\tau_x$ and $\tau_y$ are what induces tumbling, the most observable component of torque is therefore $\tau_z$, which $K_{\ell 0}$ do not affect.

The most severe effect of period on $\sigma$ is in the low-period regime ($P_\omega \lesssim 5$ hr), but in this case, the most strongly affected parameters are $K_{2m}$, which are generally known better than $K_{3m}$. The effect on the imprecise parameters $K_{3m}$ is small, except for $K_{30}$. It therefore seems as though small-period asteroids are still candidates for observation, although high-period asteroids still have better uncertainty.

\subsection{Central body oblateness}
\label{sec:scan-oblateness}

In all the above studies, we assumed a spherical planet ($J_{\ell m} = 0$ for $\ell \geq 1$). By assumption that $\mu_\mathcal{B} \gg \mu_\mathcal{A}$ (so that the asteroid orbit's focus is the centre of mass of the central body), we have $J_{1m} = 0$. The effect of central body oblateness, then, is limited to the $J_{2m}$ terms, and therefore damped by a factor of $(a_\mathcal{B} / D)^2$. We expect these parameters to have little effect on the asteroid.

Here, we define oblateness as $\epsilon = (I_z - I_x)/(\mu_\mathcal{B} R_M^2)$, where $I_{x,y,z}$ are the central body moments of inertia along the principal axes, and $I_x = I_y$. $R_M$ is the true radius of the body (not $a_\mathcal{B}$ from equation \ref{eqn:jlm}).

$J_{\ell m}$ is defined in equation \ref{eqn:jlm} with respect to the asteroid orbit, not the principal axes of the central body. However, for an equatorial orbit, the central body principal axes coincide with the asteroid orbit frame and we may express $\epsilon$ simply in terms of $J_{\ell m}$ as $\epsilon = -2J_{20}$ and $J_{22} = 0$. (Some sources such as Ref.~\cite{paterLissauer2015} take $\epsilon=J_{20}$ as their definition of $J_{20}$). For simplicity, we use this equatorial orbit case. If the orbit is non-equatorial, the other $J_{2m}$ terms will be non-zero. Since an oblate ellipsoid is mirror-symmetric around all three axes, table \ref{tab:klm-symmetries} indicates that $J_{3m}$ are all zero. The next order of precision after this oblateness approximation is therefore $J_{4m}$, damped by an additional $(a_\mathcal{B}/D)^2$ factor, and  non-ellipsoid corrections to the central body shape.

Given this conversion between $\epsilon$ and $J_{20}$, we analyze posterior uncertainty $\sigma$ of the first-order parameters as a function of $\epsilon$ across a reasonable range of central body oblatenesses based on those of Solar System planets \cite{paterLissauer2015}. These uncertainties are shown in the top pane of figure \ref{fig:scan-oblateness}, together with linear best-fitting curves for $\sigma(\epsilon)$.

\begin{figure}
  \centering
  \includegraphics[width=0.88\columnwidth]{figs/oblateness.pdf}
  \vfill
  \includegraphics[width=0.88\columnwidth]{figs/oblateness-differ.pdf}
  \caption{\textit{Top}: 1 and 2$\sigma$ confidence intervals for the first-order parameter PPDs as a function of oblateness $\epsilon$ . Linear best-fitting lines to $\sigma$ (black, dotted) are plotted. \textit{Bottom}: The difference between PPD means extracted from a zero-oblateness model and the true parameters given data with true oblateness $\epsilon_\text{true} \neq 0$. Also shown in both figures are the oblatenesses of reference Solar System bodies.}
  \label{fig:scan-oblateness}
\end{figure}

The figure demonstrates a small dependence of $\sigma$ on oblateness $\epsilon$, but the effect is far from important compared to the other factors studied above. Most Solar System bodies do not reach the oblateness necessary to significantly increase precision on the first-order parameters. The best-fitting lines match the uncertainties well, and they have slope of $(\Delta \sigma / \sigma_{\epsilon=0}) / \Delta \epsilon = -0.26$ for $\gamma_0$, $-0.94$ for $K_{22}$, and $-1.2$ for $K_{20}$. The second-order parameters $K_{3m}$ likely depend on oblateness similarly, but fitting these parameters is computationally more expensive and we do not study them here.

Note that if an encounter is executed around one of the non-Earth objects noted in figure \ref{fig:scan-oblateness}, $a_\mathcal{B}$ and $\mu_\mathcal{B}$ will change in addition to $\epsilon$. These two parameters also affect the precision of the fit parameters (appendix \ref{app:jupiter-earth}), so the figure does not show that encounters with other bodies have the same precision as encounters with Earth; only that the difference in oblateness between the two bodies is of little concern.

Given the small effect of $\epsilon$ on $K_{\ell m}$, it might be tempting to neglect the planetary oblateness when fitting $K_{\ell m}$ to data. However, the bottom pane of figure \ref{fig:scan-oblateness} demonstrates that doing so is invalid. This figure displays $K_{\ell m}$ as extracted by a fit assuming $\epsilon = 0$, but run on data generated with non-zero $\epsilon$. The difference between the PPD means and true parameters are shown. Posterior uncertainties are also shown as bands. The figure shows that even for low (Moon-scale) oblateness, the fit results are inconsistent with the true $K_{\ell m}$ values, since $\Delta K_{\ell m} = 0$ is not contained in the 2$\sigma$ band. This effect is much worse for large oblateness, growing to a difference on the order of $10^{-2}$ between the true and fit parameters for large oblateness. Therefore, accurately modelling central-body oblateness to high precision is essential for accurate estimation of fit parameters. For non-equatorial orbits, with $J_{22} \neq 0$, we also expect $J_{22}$ to affect the accuracy of the fit results to a similar degree, with the additional requirement of using the correct asteroid orbital plane.
